---
title: "User-based-Collaborative filtering"
author: "Kathirmani Sukumar"
date: "December 13, 2017"
output:
  html_document: default
  pdf_document: default
---

## Importing datasets
```{r setup}
knitr::opts_chunk$set(warning = FALSE)
library(knitr)
setwd("./")
# Use this link to download the data file
# https://github.com/skathirmani/datasets/raw/master/movies.zip
movies=read.csv("movies.csv")
ratings=read.csv("ratings.csv")
View(head(movies))
View(head(ratings))
```

 
### Data preparation 
```{r,warning=FALSE}
library(reshape2)
ratingmat <- dcast(ratings, userId~movieId, value.var = "rating", na.rm=FALSE)
dim(ratingmat)
ratingmat[1:10,1:5]
#remove userIds before coverting to rating matrix
ratingmat <- as.matrix(ratingmat[,-1]) 
View(ratingmat[1:10,1:5])
```

### Convert rating matrix into a recommenderlab sparse matrix
  * Use recommenderlab library to build recommendation engin
  
```{r, message=FALSE, warning=FALSE}
install.packages('recommenderlab')
library(recommenderlab)
ratingmat <- as(ratingmat, "realRatingMatrix")
```

### Create Recommender Model. "UBCF" stands for User-Based Collaborative Filtering

```{r, warning=FALSE}
recommender_model <- Recommender(ratingmat, method = "UBCF", param=list(method="Cosine",nn=30)) 
summary(recommender_model)
```

 * parameters passed to recommender function are method used to caluculate the similarites between the user and nn stands for Nearest Neighbors, in this example algorithm looks for 30 Nearest Neighbors.
 
### Get the top ten recommentaions for users in dataset

```{r,warning=FALSE}
Get_recomendation<-function(usr_id)
{
recom <- predict(recommender_model, ratingmat[usr_id,], n=10) 
recom_list <- as(recom, "list") #convert recommenderlab object to readable list

# getting the movie names for the list of top 10 movie ids
recom_result <- matrix(0,10)
for (i in c(1:10)){
  recom_result[i] <- as.character(movies[which(movies$movieId == as.integer(recom_list[[1]][i])),2])
}

colnames(recom_result)<-"Movies"
return(recom_result)
}

Get_recomendation(15)
```

